### 2019-09-14 ic-journal/context.tex
Stage work on multivariate section
text

\subsection{Context, guardedness and (pre)congruence}

@@

>>>>

\subsection{Context, guardedness and (pre)congruence}
\label{ss:context}

@@

Both strong bisimilarity ($\sim$) and
rooted bisimilarity ($\approx^c$) are congruence:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{congruence} \HOLConst{STRONG_EQUIV}\hfill{[STRONG_EQUIV_congruence]}
\HOLTokenTurnstile{} \HOLConst{congruence} \HOLConst{OBS_CONGR}\hfill{[OBS_CONGR_congruence]}
\end{alltt}

Although weak bisimilarity ($\approx$) is \emph{not} a congruence
  with respect to~\texttt{CONTEXT}, it is indeed substitutive
  with respect to~\texttt{GCONTEXT} 
% (or if the CCS syntax were defined with
%   only guarded sum operator \cite{sangiorgi2015equations}) 
as $\approx$ is indeed preserved by guarded sums.

Also,
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{precongruence} \HOLConst{OBS_contracts}\hfill{[OBS_contracts_precongruence]}
\end{alltt}

### 2019-09-14 ic-journal/sos.tex
text

From \hl{the point of view of HOL4, SOS rules are actually} \emph{inductive
  definitions} \hl{of the ternary} relation \HOLinline{\HOLConst{TRANS}} of type ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS} \HOLTokenTransEnd \ensuremath{\beta} \HOLTyOp{Action} \HOLTokenTransEnd \ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS} \HOLTokenTransEnd \HOLTyOp{bool}}'', generated by HOL's
\texttt{Hol_reln} function.

A useful facility exploiting the interplay
between HOL4 and Standard ML (that follows an idea by Nesi \cite{Nesi:1992ve})
 is a complex Standard ML function
  taking a CCS process and returning a theorem indicating all
  direct transitions of the process.\footnote{If the input process \hl{are
 infinite branching, the function will not terminate}.}
For instance, we know that the process $(a.\nil | \bar{a}.\nil)$ has three
possible transitions: $(a.\nil | \bar{a}.\nil) \overset{a}{\longrightarrow}
(\nil | \bar{a}.\nil)$, $(a.\nil | \bar{a}.\nil)
\overset{\bar{a}}{\longrightarrow} (a.\nil | \nil)$ and $(a.\nil | \bar{a}.\nil)
\overset{\tau}{\longrightarrow} (\nil | \nil)$.
To completely describe all possible transitions of a process, if done manually, the
following two facts must be proved: \hl{(1) there exist indeed three
  transitions mentioned above; (2) there exists no other
  transition. The first part can be easily proved by SOS rules, while
  the second part can only be proved by \texttt{TRANS\_cases}. (TODO)}

For large CCS processes it is surprisingly \hl{tedious} to manually derive
all the possible transitions and prove the non-existence of others,
\hl{which shows the usefulness of appealing to an ML function
\texttt{CCS\_TRANS\_CONV} who is designed to automate the whole process.
For instance, calling it on $(a.\nil | \bar{a}.\nil)$ the function returns the
following theorem which describes all the transitions (just one-step) of the process:}
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLTokenTransBegin\HOLFreeVar{u}\HOLTokenTransEnd \HOLFreeVar{E} \HOLSymConst{\HOLTokenEquiv{}}
   \ensuremath{(}\HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\HOLTokenDisj{}} \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\ensuremath{\mathbf{0}}}\ensuremath{)} \HOLSymConst{\HOLTokenDisj{}}
   \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\ensuremath{\mathbf{0}}}
\end{alltt}

>>>>

A useful facility exploiting the interplay
between HOL4 and Standard ML (that follows an idea by Nesi \cite{Nesi:1992ve})
 is a complex Standard ML function
  taking a CCS process and returning a theorem indicating all
  direct transitions of the process.\footnote{If the input process \hl{are
 infinite branching, the function will not terminate}.}
For instance, we know that the process $(a.\nil | \bar{a}.\nil)$ has three
possible transitions: $(a.\nil | \bar{a}.\nil) \overset{a}{\longrightarrow}
(\nil | \bar{a}.\nil)$, $(a.\nil | \bar{a}.\nil)
\overset{\bar{a}}{\longrightarrow} (a.\nil | \nil)$ and $(a.\nil | \bar{a}.\nil)
\overset{\tau}{\longrightarrow} (\nil | \nil)$.
To completely describe all possible transitions of a process, if done manually, the
following two facts must be proved: \hl{(1) there exist indeed three
  transitions mentioned above; (2) there exists no other
  transition.} For large CCS processes it is surprisingly \hl{tedious} to manually derive
all the possible transitions and prove the non-existence of others,
\hl{which shows the usefulness of appealing to an ML function
\texttt{CCS\_TRANS\_CONV} who is designed to automate the whole process.
For instance, calling it on $(a.\nil | \bar{a}.\nil)$ the function returns the
following theorem which describes all the transitions (just one-step) of the process:}
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLTokenTransBegin\HOLFreeVar{u}\HOLTokenTransEnd \HOLFreeVar{E} \HOLSymConst{\HOLTokenEquiv{}}
   \ensuremath{(}\HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\HOLTokenDisj{}} \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenOutputAct} \HOLStringLit{a} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\HOLTokenInputAct} \HOLStringLit{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\ensuremath{\mathbf{0}}}\ensuremath{)} \HOLSymConst{\HOLTokenDisj{}}
   \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{\ensuremath{\mathbf{0}}} \HOLSymConst{\ensuremath{\mid}} \HOLConst{\ensuremath{\mathbf{0}}}
\end{alltt}
